### 2025-02-22 documents/paper.tex
Add author timeseries plots.

\documentclass[11pt, a4paper, leqno]{article}
\usepackage{a4wide}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{float, afterpage, rotating, graphicx}
\usepackage{epstopdf}
\usepackage{longtable, booktabs, tabularx}
\usepackage{fancyvrb, moreverb, relsize}
\usepackage{eurosym, calc}
% \usepackage{chngcntr}
\usepackage{amsmath, amssymb, amsfonts, amsthm, bm}
\usepackage{caption}
\usepackage{mdwlist}
\usepackage{xfrac}
\usepackage{setspace}
\usepackage[dvipsnames]{xcolor}
\usepackage{subcaption}
\usepackage{minibox}
% \usepackage{pdf14} % Enable for Manuscriptcentral -- can't handle pdf 1.5
% \usepackage{endfloat} % Enable to move tables / figures to the end. Useful for some
% submissions.

\usepackage[unicode=true]{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=black,
    anchorcolor=black,
    citecolor=NavyBlue,
    filecolor=black,
    menucolor=black,
    runcolor=black,
    urlcolor=NavyBlue
}

\widowpenalty=10000
\clubpenalty=10000

\setlength{\parskip}{1ex}
\setlength{\parindent}{0ex}
\setstretch{1.5}

\begin{document}

\title{Research Proposal: Specialization Trends in Economics Research using JEL Codes \\
Text Data in Economics by Aapo Stenhammar \\
University of Bonn, Winter Term 2024/25}

\author{Felix Schmitz\footnotemark[1]}

\date{
    \today
}

\maketitle
\footnotetext[1]{Email: \href{mailto:s87fschm@uni-bonn.de}{s87fschm@uni-bonn.de}}

\begin{abstract}
    Some abstract here.
\end{abstract}

\clearpage

\section{Introduction}

\section{Data}

\subsection{Data Collection and Cleaning}

There are 6 digits in the current range without a corresponding discussion paper accessible.
The numbers are (4953, 12119, 12315, 12577, 12664, 16580).
The first discussion paper would have been from May 2010.
The later five ones fall into the time span January 2019 to November 2023.

There are 9 discussion paper entries without a corresponding discussion paper file accessible.
The numbers are (9, 18, 21, 28, 33, 56, 173, 16801, 17143).
The first nine fall into the range June 1998 to July 2000.
The later two are from February and July 2024.

These 15 entries are excluded from the analysis.

Dropped 18 entries, that have been superseded by another discussion paper.

Dropped one discussion paper each with the JEL Codes "workfromhomecommutetimesallocationoftimesavingsCOVID19" (15870), and
"labordisplacementenergytransitioncoalmines" (15581).
Dropped 16 discussion papers due to completely missing JEL Codes, these were from March 2024 to February 2025.
There were about 45 discussion papers that needed manual attention to fix the extracted JEL Codes.
Likely the JEL Codes were extracted on IZA side via Optical Character Recognition (OCR) from the PDFs, leading to confusions of e.g. "I" and "1" or "O" and "0".

\subsection{Data Collection and Cleaning}
For the analysis only data between January 2000 and December 2024 was selected.
This is mainly to ensure a sufficient number of observations per month (for the early years) and complete information on the discussion paper (for the most recent discussion papers).
This reduces the sample size by 195 observations. The final number of discussion papers analysed is 17,449.

\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{../bld/figures/fig_counts_avg_jel_codes}
    \caption{Discussion Papers \& Avg. JEL Codes per Paper Over Time}
    \label{fig:monthly_counts_and_avg_jel_codes_per_paper}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{../bld/figures/fig_author_trends}
    \caption{Average Number of Authors Over Time}
    \label{fig:author_trends}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{../bld/figures/fig_dp_counts}
    \caption{Number of Authors \& JEL Codes per Discussion Paper}
    \label{fig:dp_counts}
\end{figure}

\clearpage

\end{document}